### 2024-09-11 paper.tex
add initial structure

% PLEASE USE THIS FILE AS A TEMPLATE FOR THE PUBLICATION 
% Check file IOS-Book-Article.tex
%

\documentclass{IOS-Book-Article}     %[seceqn,secfloat,secthm]
%\usepackage{mathptmx}
%\usepackage[T1]{fontenc}
%\usepackage{times}%
%
%%%%%%%%%%% Put your definitions here

%%%%%%%%%%% End of definitions
\begin{document}
\begin{frontmatter}          % The preamble begins here.
%
%\pretitle{}
\title{}
\runningtitle{}
%\subtitle{}

% For one author:
\author{\fnms{} \snm{}}
\address{}
\runningauthor{}
%
% Two or more authors:
%\author[A]{\fnms{} \snm{}},
%\author[B]{\fnms{} \snm{}}
%\runningauthor{}
%\address[A]{}
%\address[B]{}
%
\begin{abstract}

\end{abstract}

\begin{keyword}

\end{keyword}

\end{frontmatter}

\section{Introduction}
\label{sec:intro}

\section{Background and State of the Art}
\label{sec:sota}

- DPV
- DCMI 
- PROV
- DCAT
- ODRL

\section{Rights Exercise and Management}
\label{sec:rights}

\begin{itemize}
    \item[Section 3.1.] Information about the existence of rights
    \item[Section 3.2.] Justifications
    \item[Section 3.3.] Notifications related to rights activities
    \item[Section 3.4.] Records of rights activities
    \item[Section 3.5.] GDPR-related rights requests as machine-executable policies
\end{itemize}

\section{Conclusions and Future Work}
\label{sec:conclusions}

\bibliography{ref}
\bibliographystyle{vancouver}

\end{document}